The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate

### 2020-11-16 VegOzone.tex
First words about results.

In terms of $\mathrm{AOT_{40}}$, the threshold for protection of vegetation of $3000\,\unit{ppb}$ (integrated over the growing season) given by the EU directive has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019}. {\bf TODO: come back to this later on in the discussions(?) - it has not been breached at Svanvik in northern Fennoscandia. Later supplement figure \ref{fig:fennoscandic_aot40}}

@@

As pointed out by \citet{ACP:Bueker2012}, a very important factor for deriving $\text{POD}_y$ is the soil texture or how much water the soil can hold. As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture at $60\,\unit{cm}$ depth.

@@

{\bf TODO: Model results for natural and semi-natural vegetation (birch, pine, grassland)}

\begin{figure*}[t]
  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3_Birch}
  \caption{.}
  \label{fig:pody_birch}
\end{figure*}

\begin{figure*}[t]
  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3_Norway_spruce}
  \caption{.}
  \label{fig:pody_norway_spruce}
\end{figure*}

\begin{figure*}[t]
  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3_Perennial_grass}
  \caption{.}
  \label{fig:pody_grass}
\end{figure*}

%\begin{table}[t]
%\caption{TEXT}
%\begin{tabular}{column = lcr}
%\tophline
%
%\middlehline
%
%\bottomhline
%\end{tabular}
%\belowtable{} % Table Footnotes
%\end{table}

@@

\begin{figure*}[t]
  \includegraphics[width=12cm]{modis_Psn}
  \caption{Estimated begin (BGS) and end (EGS) of growing season for coniferous trees from MODIS Aqua/Terra net photosynthesis (PSN) product. A $1\times 1\,\unit{km}$ area around Svanhovd was selected. Daily averaged data for both 2018 and 2019 has been fitted with a quadratic polynomial function. The numerically computed root yields: BGS 122/106 day of year and EGS 261/274 day of year for 2018/2019, respectively.}
  \label{fig:modis_Psn}
\end{figure*}

>>>>

In terms of $\mathrm{AOT_{40}}$, the threshold for protection of vegetation of $3000\,\unit{ppb}$ (integrated over the growing season) given by the EU directive has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019}. {\bf TODO: come back to this later on in the discussions(?) - it has not been breached at Svanvik in northern Fennoscandia.}

@@

As pointed out by \citet{ACP:Bueker2012}, a very important factor in deriving a proper $\text{POD}_y$ is the soil texture or how much water the soil can hold. As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture at $60\,\unit{cm}$ depth.

The estimated total biomass reduction in \unit{\%} due to \chem{O_3} following \citet{ICP:MappingManual2017} is shown in Table~\ref{tab:biomass_reduction}.

@@

We have modeled $\text{POD}_1$ for three natural and semi-natural vegetation PFTs, birch, Norway spruce, and perennial grassland. The results are shown comprehensively in Fig.~\ref{fig:pody} as a function of day of year (doy) for March--October. \chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$. Due to expected non linearities in stomatal conductance, the climatological year represents rather an optimal year then a true climatology in this case. To establish a true climatology in $\text{POD}_y$, it would therefore be necessary to run the DO3SE model for 10--20 years with true start and end of growing season for each PFT and the respective meteorological and ozone forcing.

As expected $\text{POD}_y$ was higher in 2018 than in 2019 for all PFTs.

\begin{figure*}[t]
  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3}
  \caption{$\text{POD}_y$ modeling results from DO3SE model for three different PFTs over day of year (doy), March--October. A flux threashold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen. \chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$. Due to non linearities in stomatal conductance, the climatological year represents rather an optimal year then a true climatology. (a)--(c) Birch; (d)--(f) Norway spruce; (g)--(i) perennial grassland. From left to right: 2018, 2019, and climatological year.}
  \label{fig:pody}
\end{figure*}

\begin{table}[t]
  \caption{Estimated total biomass reduction in \unit{\%} due to \chem{O_3} following \citet{ICP:MappingManual2017}.}
  \label{tab:biomass_reduction}
\begin{tabular}{lccrl}
\tophline
Year & \multicolumn{4}{c}{PFT}\\
& Birch & Norway spruce & \multicolumn{2}{c}{Perennial grassland}\\
\middlehline
2018 & 11.99 & 2.34 & 8.06 & (10.51)$^*$\\
2019 & 10.07 & 2.26 & 6.43 & (7.90)$^*$\\
clim. & 13.81 & 2.58 & 6.67 & (8.29)$^*$\\
\bottomhline
\end{tabular}
\belowtable{$^*$ above ground bioamass reduction} % Table Footnotes
\end{table}

@@

\begin{figure*}[t]
  \includegraphics[width=12cm]{modis_Psn}
  \caption{Estimated begin (BGS) and end (EGS) of growing season for coniferous trees from MODIS Aqua/Terra net photosynthesis (PSN) product. A $1\times 1\,\unit{km}$ area around Svanhovd was selected. Daily averaged data for both 2018 and 2019 has been fitted with a quadratic polynomial function. The numerically computed root yields: BGS 122/106 day of year (doy) and EGS 261/274 day of year for 2018/2019, respectively.}
  \label{fig:modis_Psn}
\end{figure*}
